Boosting Higgs Pair Production in the $b\bar{b}b\bar{b}$ Channel at the LHC with Multivariate Techniques

### 2015-07-14 paper/sec-mva.tex
work on paper

As can be seen from Fig.~\ref{fig:nev2}, in the boosted category
event with a cut in the ANN output as hard as $y_{\rm cut}\simeq 0.9$
we are still left
with around 100 events at the HL-LHC, which is a reasonable number,
which the background being reduced substantially to around
500 events.
%
A similar cut would work fine for the intermediate category, though here
the backgrounds are much larger and thus the expected signal significance smaller.
%
For the resolved case, the signal significance would be only slightly
improved for any value of the cut.

A very useful property of MVAs such as the one used in this work
is that they provide direct  physical insight about which of the
input variables contribute the most to the signal over background
separation.
%
In the case of ANNs, this can be quantified by computing the sum
of the absolute value of al the weights connected to a given
input neuron $i$, that is
\be
\label{eq:totweight}
\omega^{\rm tot}_i \equiv \sum_{k=1}^{n^{(2)}} \Big|\omega^{(2)}_{ki}\Big| \, ,
\qquad i=1,\ldots,N_{\rm var} \, ,
\ee
with $\omega^{(2)}_{ki}$ the value of the weight connecting
the $k$-th neutron of the second layer with the $i$-th neuron of
the first (input) layer, and $n^{(2)}=5$ the number of
neurons in the second layer.
%
The input variables with a larger value of $\omega^{\rm tot}_i$ will those
be the ones that play a dominant role in enhancing the signal
discrimination using the MVA.

%
In Fig.~\ref{fig:nnweights} we show
the distribution of the total associated weight,
Eq.~(\ref{eq:totweight}) for each of the $N_{\rm var}$ input
variables of the resolved (left plot) and boosted (right plot) categories.
%
The notation for the various kinematic variables is the same
as in Sect.~\ref{sec:input}.
%
The scale of the $y$ axis is arbitrary, and the important information
is contained in the relative strengths of the total associated weight
for each of the input variables.

%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[t]
\begin{center}
\includegraphics[width=0.48\textwidth]{plots/res_wgthist.pdf}
\includegraphics[width=0.48\textwidth]{plots/boosted_wgthist.pdf}
\caption{\small
Distribution of the total associated weight,
Eq.~(\ref{eq:totweight}) for each of the $N_{\rm var}$ input
variables of the resolved (left plot) and boosted (right plot) categories.
%
See Sect.~\ref{sec:input} for a description of each of the input
variables to the MVA in the two cases.
}
\label{fig:nnweights}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%

The results of Fig.~\ref{fig:nnweights} clearly illustrate which variables
are more discriminative between the signal and the background.
%
Let's begin discussing the resolved analysis.
%
In this category, the variables that lead to
a higher discrimination power
are the $p_T$ of the two reconstructed Higgs, and then on a similar
footing other variables like the $p_T$ of the individual jets
and the mass of the di-Higgs system.
%
The only variable which seems to contain to information at all
is the $\Delta \phi_{hh}$ separation between the two
Higgs candidates.

>>>>

As can be seen from Fig.~\ref{fig:nev2}, in the boosted category
event with a cut in the ANN output as hard as $y_{\rm cut}\simeq 0.9$
we are still left
with around 100 events at the HL-LHC, which is a reasonable number,
which the background being reduced substantially to around
1000 events.
%
A similar cut would work fine for the intermediate category, though here
the backgrounds are much larger and thus the expected signal significance smaller.
%
For the resolved case, the signal significance would be only slightly
improved for any value of the cut.

A useful property of MVAs such as the one used in this work
is that they provide direct  physical insight about which of the
input variables contribute the most to the separation of
signal over background.
%
In the case of ANNs, this can be quantified by computing the sum
of the absolute value of al the weights connected to a given
input neuron $i$, that is
\be
\label{eq:totweight}
\omega^{\rm tot}_i \equiv \sum_{k=1}^{n^{(2)}} \Big|\omega^{(2)}_{ki}\Big| \, ,
\qquad i=1,\ldots,N_{\rm var} \, ,
\ee
with $\omega^{(2)}_{ki}$ the value of the weight connecting
the $k$-th neutron of the second layer with the $i$-th neuron of
the first (input) layer, and $n^{(2)}=5$ the number of
neurons in the second layer.
%
The input variables with a larger value of $\omega^{\rm tot}_i$ will those
be the ones that play a dominant role in enhancing the signal
discrimination using the MVA.

%
In Fig.~\ref{fig:nnweights} we show
the distribution of the total associated weight,
Eq.~(\ref{eq:totweight}) for each of the $N_{\rm var}$ input
variables of the resolved (left plot) and boosted (right plot) categories.
%
The notation for the various kinematic variables is the same
as in Sect.~\ref{sec:input}.
%
The important information
is contained in the relative strengths of the total associated weight
for each of the input variables.
%
The results of Fig.~\ref{fig:nnweights} illustrate which variables
are more discriminative between the signal and the background.
%
For instance, in the 
resolved category, the variables that lead to
a higher discrimination power
are the $p_T$ of the two reconstructed Higgs candidates, and then on a similar
footing other variables like the $p_T$ of the individual jets
and the mass of the di-Higgs system.
%
The only variable which seems to contain to information at all
is the $\Delta \phi_{hh}$ separation between the two
Higgs candidates.

%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[t]
\begin{center}
\includegraphics[width=0.49\textwidth]{plots/res_wgthist.pdf}
\includegraphics[width=0.49\textwidth]{plots/boosted_wgthist.pdf}
\vspace{-0.5cm}
\caption{\small
Distribution of the total associated weight,
Eq.~(\ref{eq:totweight}) for each of the $N_{\rm var}$ input
variables of the resolved (left plot) and boosted (right plot) categories.
%
See Sect.~\ref{sec:input} for a description of each of the input
variables to the MVA.
}
\label{fig:nnweights}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%
